---
title: "Lesson 2 Prep"
author: "Samantha Chinn"
date: "1/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(faraway)
```

```{r}
gavote
```

```{r}
str(gavote)
```

```{r}
summary(gavote)
```

Create new variable undercount:
```{r}
gavote$undercount <- (gavote$ballots-gavote$votes) / gavote$ballots
summary(gavote$undercount)
```

Overall relative undercount:
with command removes the necessity for gavote$ every time
```{r}
with(gavote, sum(ballots-votes)/sum(ballots))
```

Histograms:
```{r}
hist(gavote$undercount, main="Undercount", xlab="Percent Undercount")
```

```{r}
plot(density(gavote$undercount),main="Undercount")
rug(gavote$undercount)
```

```{r}
pie(table(gavote$equip),col=gray(0:4/4))
```

Pareto Chart - bar plot with categoreis in descending order of frequency.
```{r}
barplot(sort(table(gavote$equip), decreasing=TRUE), las=2)
```

```{r}
gavote$pergore <- gavote$gore/gavote$votes
plot(pergore ~ perAA, gavote, xlab="Proportion African American", ylab="Proportion for Gore")
```

Side-by-side boxplots:
```{r}
plot(undercount ~ equip, gavote, xlab="", las=3)
```

```{r}
xtabs(~ atlanta + rural, gavote)
```

```{r}
names(gavote)
```

```{r}
names(gavote)[4] <- "usage"
```

```{r}
nix <- c(3,10,11,12)
cor(gavote[,nix])
```

Fitting a Linear Model:
```{r}
lmod <- lm(undercount ~ pergore + perAA, gavote)
```

```{r}
coef(lmod)
```

```{r}
predict(lmod)
```

```{r}
residuals(lmod)
```

```{r}
deviance(lmod)
```

```{r}
df.residual(lmod)
```

```{r}
nrow(gavote)-length(coef(lmod))
```

```{r}
sqrt(deviance(lmod)/df.residual(lmod))
```

```{r}
lmodsum <- summary(lmod)
lmodsum$sigma
```

```{r}
lmodsum$r.squared
```

```{r}
cor(predict(lmod),gavote$undercount)^2
```

```{r}
lmodsum$adj.r.squared
```

```{r}
summary(lmod)
```

```{r}
library(faraway)
sumary(lmod)
```

Interpretation:
```{r}
gavote$cpergore <- gavote$pergore - mean(gavote$pergore)
gavote$cperAA <- gavote$perAA - mean(gavote$perAA)
lmodi <- lm(undercount ~ cperAA+cpergore*usage+equip, gavote)
sumary(lmodi)
```

Hypothesis Testing:
```{r}
anova(lmod,lmodi)
```

Comparison of all models with one predictor less than the larger model:
```{r}
drop1(lmodi,test="F")
```

Confidence Intervals:
```{r}
confint(lmodi)
```

Diagnostics:
```{r}
plot(lmodi)
```

```{r}
gavote[cooks.distance(lmod) > 0.1,]
```

Half-normal plot:
```{r}
halfnorm(hatvalues(lmodi))
```

```{r}
gavote[hatvalues(lmodi)>0.3,]
```

```{r}
termplot(lmodi,partial=TRUE,terms=1)
```

Robust Regression:
```{r}
library(MASS)
rlmodi <- rlm(undercount ~ cperAA+cpergore*usage+equip, gavote)
summary(rlmodi)
```

Weight:
```{r}
wlmodi <- lm(undercount ~ cperAA+cpergore*usage+equip, gavote, weights=ballots)
```

Transformation:
```{r}
plmodi <- lm(undercount ~ poly(cperAA,4)+cpergore*usage+equip,gavote)
summary(plmodi)
```

```{r}
termplot(plmodi, partial=TRUE,terms=1)
```

